Portfolios of Subgraph Isomorphism Algorithms

### 2015-12-06 paper/paper.tex
Rework and deduplicate a bit

\section{Background}

\todo[inline]{Some background on subgraph isomorphism algorithms, maybe move
some stuff from the intro here.}

@@

Here, we consider the case where exactly one algorithm is selected for solving the problem. One of
the most prominent and successful systems that employs this approach is
SATzilla~\cite{xu_satzilla_2008}, which defined the state of the art in SAT solving for a number of
years. Other application areas include constraint solving~\cite{omahony_using_2008}, the travelling
salesperson problem~\cite{kotthoff_improving_2015}, and AI planning~\cite{seipp_learning_2012}.

The interested reader is referred to a recent survey~\cite{kotthoff_algorithm_2014} for additional
information on algorithm selection.

\section{Algorithms for subgraph isomorphism}

We consider the following exact algorithms: VF2 \cite{Cordella:2004}, which is a simple backtracking
search algorithm, especially fast on trivially satisfiable instances; LAD \cite{Solnon:2010}, which
does very strong reasoning based on local all different redundant constraints; and the Glasgow
algorithm \cite{McCreesh:2015}, which does an expensive preprocessing pass once, followed by weaker
reasoning during search.

@@

We introduce two new variants of LAD. The first variant is a weaker LAD filtering (called LAD
preprocessing) which is applied once, without performing a backtracking search, and detects
inconsistencies on many instances: For each pattern node $u$, we check that there exists at least
one target node $v$ such that for each neighbor $u'$ of $u$ there exists a different neighbor $v'$
of $v$ such that the degree of $u'$ is smaller than or equal to the degree of $v'$. LAD
preprocessing is an incomplete algorithm as it can only detect inconsistency of some instances: When
it does not detect inconsistency, the instance may either be consistent or not. The time complexity
of LAD preprocessing is ${\cal O}(n(n'+p'))$ where $n$ and $n'$ are the number of pattern and target
nodes, respectively, and $p$ is the number of target edges.

@@

\section{Benchmark}

We consider 5725 instances which are available in a simple text
format\footnote{\url{http://liris.cnrs.fr/csolnon/SIP.html}}. These instances are grouped into 12
classes.

@@

\noindent Randomly generated instances with particular properties:

\begin{itemize}
    \item Bounded-degree random graphs (family 3, 4) \cite{GraphDatabase}
    \item Random 4d mesh graphs (family 5, 6) \cite{GraphDatabase}
    \item Random graphs (family 7) \cite{GraphDatabase}
    \item Randomly generated scale-free graphs. (family 1) \cite{Solnon:2010}
    \item Phase transition (family 12)
\end{itemize}

We consider a total of 5725 instances.

\subsubsection{Features}

\todo[inline]{Describe features here.}

This gives us a total of 30 features.

@@

\section{Results}

@@

\section{Experimental comparison}

Our runtimes were measured on machines with Intel Xeon E5-2640 v2 CPUs and 64GBytes RAM, running
Scientific Linux 6.5. We used the C++ implementation of the Glasgow algorithm \cite{McCreesh:2015},
the C implementation of LAD \cite{Solnon:2010}, and the VFLib C implementation of VF2
\cite{Cordella:2004}. Software was compiled using GCC 4.9. Each problem instance was run with a
timeout of $10^8$ milliseconds (a little over a day).

@@

Table \ref{expClass} shows us that we cannot reasonably split things up based upon which class
instances are comin from, unlike SAT. For all classes, there are always at least 2 algorithms which
are the best for at least one instance of the class. In particular, for classes 2 and 3, each
algorithm is the best for at least one instance (except Glasgow 4 for Class 2).

@@

\section{Algorithm selection approach}

Describe features + selection approach.

@@

>>>>

\section{Per-Instance Algorithm Selection}

@@

Here, we consider the case where exactly one algorithm is selected for solving the problem. One of
the most prominent and successful systems that employs this approach is
SATzilla~\cite{xu_satzilla_2008}, which defined the state of the art in SAT solving for a number of
years. Other application areas include constraint solving~\cite{omahony_using_2008}, the travelling
salesperson problem~\cite{kotthoff_improving_2015}, and AI planning~\cite{seipp_learning_2012}.
The interested reader is referred to a recent survey~\cite{kotthoff_algorithm_2014} for additional
information on algorithm selection.

\subsection{Subgraph isomorphism algorithms}

We consider the following exact subgraph isomorphism
algorithms: VF2 \cite{Cordella:2004}, which is a simple backtracking search algorithm, especially
fast on trivially satisfiable instances; LAD \cite{Solnon:2010}, which does very strong reasoning
based on local all different redundant constraints; and the Glasgow algorithm \cite{McCreesh:2015},
which does an expensive preprocessing pass once, followed by weaker reasoning during search.

@@

We also introduce two new variants of LAD. The first variant is a weaker LAD filtering (called LAD
preprocessing) which is applied once, without performing a backtracking search, and detects
inconsistencies on many instances: For each pattern node $u$, we check that there exists at least
one target node $v$ such that for each neighbor $u'$ of $u$ there exists a different neighbor $v'$
of $v$ such that the degree of $u'$ is smaller than or equal to the degree of $v'$. LAD
preprocessing is an incomplete algorithm as it can only detect inconsistency of some instances: when
it does not detect inconsistency, the instance may either be consistent or not. The time complexity
of LAD preprocessing is ${\cal O}(n(n'+p'))$ where $n$ and $n'$ are the number of pattern and target
nodes, respectively, and $p$ is the number of target edges.

@@

\subsection{Problem instances}

We consider 5725 instances, which are available in a simple text
format\footnote{\url{http://liris.cnrs.fr/csolnon/SIP.html}}. These instances are grouped into 12
classes.

@@

\subsection{Features}

For both the pattern and the target graph, we consider the number of vertices, edges and loops, the
mean and maximum degree, the density, whether or not every vertex has same degree, whether or not
the graph is connected, the mean and maximum distances between all pairs of vertices, and the
proportion of vertex pairs which are distance at least 2, 3 and 4 apart. We also consider the number
of variable-value assignments removed by LAD preprocessing in absolute terms, as a percentage, and
the minimum and maximum on a per-variable basis.  This gives us a total of 30 features.

\subsection{Experimental Setup}

We measured runtimes on machines with Intel Xeon E5-2640 v2 CPUs and 64GBytes RAM, running
Scientific Linux 6.5. We used the C++ implementation of the Glasgow algorithm \cite{McCreesh:2015},
the C implementation of LAD \cite{Solnon:2010}, and the VFLib C implementation of VF2
\cite{Cordella:2004}. Software was compiled using GCC 4.9. Each problem instance was run with a
timeout of $10^8$ milliseconds (a little over a day).

@@

\section{Experimental Results}

@@

@@

Table \ref{expClass} shows us that we cannot reasonably split things up based upon which class
instances are coming from, unlike SAT. For all classes, there are always at least 2 algorithms which
are the best for at least one instance of the class. In particular, for classes 2 and 3, each
algorithm is the best for at least one instance (except Glasgow 4 for Class 2).

@@

@@

Mention parallel, directed, labelled in here briefly.
